### 2016-03-16 main.tex
Make the paper shorter

\documentclass[twocolumn]{article}
%\documentclass{article}

\usepackage[utf8]{inputenc}

\usepackage{graphicx} 
\usepackage{subfigure}
\usepackage{paralist}
\usepackage{amsfonts}
\usepackage{hyperref}

\usepackage{url}
\usepackage{booktabs}

\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{tikz}
\usetikzlibrary{positioning, calc}

\usepackage[draft,nomargin,footnote]{fixme}

\graphicspath{{figs/}}

\usepackage{xspace}
\newcommand{\eg}{\textit{e.g.}\xspace}
\newcommand{\etal}{\textit{et al.}\xspace}
\newcommand{\ie}{\textit{i.e.}\xspace}
\newcommand{\etc}{\textit{etc.}\xspace}
\newcommand{\vs}{\textit{vs.}\xspace}
\newcommand{\resp}{\textit{resp.}\xspace}

\title{Learning by Teaching a Robot:\\The Case of Handwriting}

\author{Séverin Lemaignan$^1$, Alexis Jacq$^{1,2}$, Deanna Hood$^1$,\\Fernando
    Garcia$^1$, Ana Paiva$^2$, Pierre Dillenbourg$^1$ \\
$^1$CHILI Lab, École Polytechnique Fédérale de Lausanne, Switzerland,\\
$^2$Instituto Superior Técnico, University of Lisbon, Portugal}

\begin{document}
\maketitle

%\begin{abstract}
%
%    Robots for education are not limited to support ICT teaching, and they are
%    indeed finding new roles in the classroom. This article reports on such a
%    new paradigm for educative robots, that involves \emph{learning by teaching}
%    and strong \emph{social engagement} to help children struggling with
%    handwriting. Our system relies on machine-learning and child-robot
%    interaction with a small humanoid robot, and we present several real-world
%    studies in schools and with occupational therapists that led us to promising
%    initial results.
%
%\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A Different Paradigm for Educative Robots}

Thomas is five and half, and has been diagnosed with visuo-constructive
deficits. He is under the care of an occupational therapist, and tries to
work around his inability to draw letters in a consistent manner. Vincent is six
and struggles at school with his poor handwriting and even poorer
self-confidence.\footnote{All children's names have been changed.}

While Thomas is lively and always quick at shifting his attention from one
activity to another, Vincent is shy and poised. Two very different children,
facing however the same difficulty to write in a legible manner. And, hidden
beyond this impaired skill, psycho-social difficulties arise: they underperform
at school, Thomas has to go for follow-up visits every week, they both live
under the label ``requires special care''. This is a source of anxiety for the
children and for their parents alike.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{henry}
    \caption{\small Thomas teaching a {\sc Nao} robot how to write numbers, with 
    the help of an occupational therapist.}
    \label{fig:henry}
\end{figure}

Remediations for handwriting difficulties traditionally involve long
interventions (at least 10 weeks~\cite{Hoy2011}), essentially consisting of
handwriting training with occupational therapists, and primarily addressing the
\emph{motor} deficits.  Improvements in self-confidence and anxiety occur (at
best) as a side-effect of the child improving their handwriting skills and,
consequently, improving their performance at school.

We present in this article a new take on this educative challenge: a remediation
procedure that involves a ``bad writer'' robot that is \emph{taught} by the
child. By building on the \emph{learning by teaching} paradigm, not only does the
child practise handwriting but, as they take on the role of the teacher, they
also positively reinforce their self-esteem and motivation: their social role
shifts from the ``underperformer'' to ``the one who knows and teaches''. And by
relying on a robot, we can tailor the exercises and the learning curve to each
child's needs individually, as we will show in this article.

\subsection*{Learning by Teaching Handwriting}

The \emph{learning by teaching} paradigm, which engages the student in the act
of teaching another, has been shown to produce motivational, meta-cognitive, and
educational benefits in a range of disciplines~\cite{Rohrbeck2003}. The
application of this paradigm to handwriting intervention remains, however,
unexplored. One reason for this may be due to the requirement of an
appropriately unskilled peer for the child to tutor: this may indeed prove
difficult if the child is the lowest performer in the class.  In some cases, it
may be appropriate for a peer or teacher to simulate a na\"ive learner for the
child to teach. For handwriting however, where one's skill level is visually
evident, this acting is likely to be rapidly detected. This motivates the use of
an artificial teachable agent which can be configured for a variety of skill
levels, and for which children do not have preconceptions about its handwriting
ability.

Robots have been used as teachers or social partners to promote children's
learning in a range of contexts, most commonly related to language
skills~\cite{han2010robot}, and less often to physical skills (such as
calligraphy~\cite{Matsui2013}). Looking at the converse (humans \emph{teaching}
robots), Werfel notes in~\cite{Werfel2014} that most of the work focuses on the
robot's benefits (in terms of language~\cite{Saunders2010} or
physical~\cite{Mulling2013} skills, for example) rather than the learning
experienced by the human tutor themselves.  Our work concentrates on this latter
aspect: by demonstrating handwriting to a robot, we aim at improving the
\emph{child's} performance.

A robotic learning agent which employs the learning by teaching paradigm has
previously been developed by Tanaka and Matsuzoe~\cite{Tanaka2012}. In their
system, children learn vocabulary by teaching the {\sc nao} robot to act out
verbs. The robot is tele-operated (Wizard-of-Oz) and mimics the actions that the
children teach it, but with no long-term memory or learning algorithm in place.
Our project significantly extends this line of work in two ways. First, by
investigating the context of children's acquisition of a challenging physical
skill (handwriting), and second by proposing a robotic partner which is fully
autonomous in its learning.

\subsection*{Agency and Commitment}

We also investigate here a particular role for a robot in the education of
handwriting: not only is the robot actively performing the activity by drawing
letters, but it does so in a way that engages the child in a very specific
social role. The child is the teacher in this relationship and the robot is the
learner: the child is to engage in a (meta-) cognitive relationship with the
robot to try to understand why the robot fails and how to best help it.  Here,
the robot is more than just an activity facilitator or orchestrator -- its
physical presence and embodiment induce agency and anthropomorphizing, and
cognitively engage the child into the learning activity (be it consciously or
not).

The commitment of the child into the interaction builds on a psychological effect
known as the ``protégé effect''~\cite{Chase2009}: the teacher feels responsible
for his student, commits to the student's success and possibly experiences
student's failure as his own failure to teach. Teachable computer-based agents
have previously been used to encourage this ``protégé effect'', wherein students
invest more effort into learning when it is for a teachable agent than for
themselves~\cite{Chase2009}.  We rely on this cognitive mechanism to reinforce
the child's commitment into the robot-mediated handwriting activity, and we
indeed show sustained child-robot engagement over extended periods of time
(several hours spread over a month).

For these two reasons, our approach is to be distinguished from previous works
in educational robotics. Most of these do not consider the agency induced by the
robot beyond its motivational aspect (playing with an interactive, partially
autonomous device induces excitement -- at
least, on the short-term).  In our case, the
role of agency is stronger: it induces meta-cognition (``I am interacting with
an agent, so I need to reflect on how to best teach him'') which is beneficial
for the learning process; it also induces a protégé effect (``I want my
robot-agent to succeed!'') which supports the commitment of the child into the
interaction, also for longer periods of time.

\vspace{2em}
These initial considerations can be turned into a set of research questions: how
to develop a fully autonomous robotic system that produces believable
handwriting, foster interaction and adequately learn from children's demonstration?
Can such a system be accepted by children and practitioners alike, and
eventually be integrated in existing educative and therapeutic practises?
Can we build and evidence interactions that sustains children's engagement over extended
periods of time? And finally, is this novel approach effective? Do we evidence
actual handwriting improvements by children facing difficulties?

We attempt to provide initial insights on these questions in the remaining
sections of the article. We first provide an overview of the robotic system and
the interaction it induces in Section~\ref{impl}. We present there the
machine-learning techniques that allow the robot to learn from the children, as
well as the actual implementation of the system on a {\sc nao} robot.

We then present and report on the field experiments that we have conducted over
the last two years, including four studies at schools
(Section~\ref{studies_school}), one longer experiment with eight children in an
occupational therapy clinic (Section~\ref{study_normandy}), and two one-month
long case studies (Sections~\ref{vincent} and~\ref{thomas}). While the focus of
the school experiments was mostly the technical validation of the system and
data acquisition, the three other experiments involved children with actual
deficits, and gave us initial insights into the relevance and effectiveness of
our approach.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implementation of the Interaction}
\label{impl}

Figure~\ref{experimental_setup} illustrates our general experimental setup: a
face-to-face child-robot interaction with an autonomous Aldebran's {\sc nao}
robot. A tactile tablet is used for both the robot and the
child to write. During a typical round, the child requests the robot to write
something (a single letter, a number or a full word), and pushes the tablet
towards the robot. The robot writes on the tablet by gesturing the writing in
the air,  the letters being actually drawn by the tablet application itself. The
child then pulls back the tablet, corrects the robot's attempt by writing
herself on top of or next to the robot's writing (see Figure~\ref{fig:diego}),
and sends her demonstration to the robot by pressing a small button on
the tablet. The robot learns from this demonstration and tries again.

Since the children are assumed to take on the role of the teachers, we had to
ensure they would be able to manage by themselves the turn-taking and the
overall progression of the activity (moving forwards to the next letter or
word). In our design, the turn-taking relies on the robot prompting for feedback
once it is done with its writing (through simple sentences like ``What do you
think?''), and pressing on a small robot icon on the tablet once the child has
finished correcting. In our experiments, once introduced by the
experimenter, both steps were easy to grasp for the children.

\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{experimental_setup}
    \caption{\small Our experimental setup: the robot writes on a tactile tablet, the child then
        corrects the robot by overwriting its letters with a stylus. An adult
        (either a therapist or an experimenter) guides the work
        (prompting, turn taking, etc.). For some studies, a second tablet and an
        additional camera (dashed) are employed.}

    \label{experimental_setup}
\end{figure}

Implementing such a system raises several challenges: first, the acquisition,
analysis and learning from hand-written demonstration, which lays at the core of
the our approach, necessitates the development of several algorithms for the
robot to generate initial bad writing and to respond in an adequate manner,
showing visible (but not too quick) writing improvements.

Then, the actual implementation on the robot requires the coordination of
several modules (from performing gestures and acquiring the user's input to the
state machine implementing the high-level behavior), spread over several
devices (the robot itself, one laptop and up to four tactile tablets for some of
the studies that we conducted). We relied on ROS to ensure the synchronization
and communication between these modules.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Generating and Learning Letters}

Since our application is about teaching a robot to write, generating (initially
bad) letters and learning from writing demonstrations is a core aspect of the
project.

The main insight for both the generation and the learning of letters is to
reason about the shape of letters in their eigenspace, instead of the natural
cartesian space. The eigenspace of each letter is spanned by the first $n$
eigenvectors (in our experiments, $3 < n < 6$) of the covariance matrix
generated from a standard dataset of adult letters (the UJI Pen Characters 2
dataset~\cite{Llorens2008}).  Our method, based on a Principle Component
Analysis (PCA), is presented in~\cite{hood2015when}.

New letters are generated by varying the weightings in linear combinations of
these eigenvectors, with distortions that are actually plausible handwriting errors: 
they are exaggerations of variations of writing styles that naturally occur amongst
adult writers.  Figure~\ref{fig:sampleLetters} shows examples of deformed ``g''
generated with such a technique.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{cowriter-g}
    \caption{\small \label{fig:sampleLetters} \textbf{Generating bad letters}:
        effect of independently varying the weightings (columns) of the first five eigenvectors 
        (rows) of the shape model of ``g''. Examples A, B, C, D illustrate how the
        PCA-based approach allows for the \emph{automatic} generation of letters whose
        errors are typical of children handwriting: A has too large a bottom
        loop, B has a wide top loop, the bottom loop of C is not correctly
        closed, the top loop of D is not closed, etc.}

\end{figure}

The same shape model can also be used to \emph{classify} demonstrations,
\emph{assess} their quality and \emph{learn} from them. Figure~\ref{fig:h} shows
for example nine allographs of ``h'' written by a 6 years old child, along with
the reference shape. By projecting each of the
demonstrations onto the eigenspace of ``h'' (Figure~\ref{fig:h}b), we observe
that:

\begin{itemize}
    \item the different allographs can by clustered (with a $k$-means or
        mean-shift algorithm) by their visual styles,

    \item we can compute a euclidian distance to the reference letter to assess
        the topological proximity of the demonstration with the expected letter,
        thus providing a quantitative metric of writing performance.

\end{itemize}

\begin{figure*}[ht!]
    \centering
    \subfigure[Nine allographs of the cursive ``h'', next to the reference]{
        \includegraphics[width=0.45\linewidth]{h}
    }
    \subfigure[The same samples, normalized and projected in the eigenspace spanned by the
        first 3 eigenvectors: clusters arise, that actually match writing styles.]{
        \includegraphics[width=0.45\linewidth]{eigenspace-uniformization}
    }

    \caption{\small Projecting demonstrated letters onto the eigenspace
    generated from the reference dataset effectively clusters the samples
    according to their topological similarity. Allographs that are similar to
    the reference are close to it in the eigenspace.}
    \label{fig:h}
\end{figure*}

The algorithm for machine-learning becomes then a simple matter of converging at
a specific pace towards the child's demonstration in the eigenspace.
Figure~\ref{learning_6_demos} (p.~\pageref{learning_6_demos}) illustrates the
process with a complete learning cycle of the number ``6''.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Robotic Implementation}

Our system is embodied in an Aldebaran {\sc nao} (V4 or V5, depending on the
study) humanoid robot. This choice is motivated by its approachable
design~\cite{Gouaillier2008}, its size (58cm) and inherently safe structure
(lightweight plastic) making it suitable for close interaction with children,
its low price (making it closer to what schools may afford in the coming years)
and finally its ease of deployment in the field.

Robotic handwriting requires precise closed-loop control of the arm and hand
motion. Because of the limited fine motor skills possible with such an
affordable robot, in addition to the absence of force feedback, we have opted
for \emph{simulated handwriting}: the robot draws letters in the air, and the
actual writing is displayed on a synchronized tablet.

The overall architecture of the system is therefore
spread over several devices: the {\sc nao} robot itself, that we address via
both a ROS API and the Aldebaran-provided NaoQI API, one to
four Android tablets (the main tablet is used to draw the robot's letter and to
acquire the children's demonstrations; more tablets have been used in some
studies, either to let the child input words to be written, or for the
experimenter to qualitatively annotate the interaction in a synchronized
fashion), and a central laptop running the machine learning algorithms, the
robot's handwriting gesture generation (based on the NaoQI inverse kinematics
library) and the high-level control of the activity (relying on {\sc
pyRobots}~\cite{lemaignan2015pyrobots} and a custom finite state machine).

Since the system does not actually require any CPU-intensive processes, the laptop
can be removed and the whole logic run on the robot. Due to the relative
difficulty to deploy and debug ROS nodes directly on the robot, the laptop
remains however convenient during the development phase and we kept it during
our experiments.

Most of the nodes are written in Python, and the whole source code of the
project is available online\footnote{The primary repository is
\url{https://github.com/chili-epfl/cowriter_letter_learning}.}. The details of
the technical implementation are available in~\cite{hood2015when}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Field Studies}

The system has been deployed and tested in several situations: in three
different schools (more than 70 children aged 5 to 8; relatively short duration
interactions), in an occupational therapy clinic (8 children, each
interacting several hours with the system), and during two case studies that
each lasted several weeks.
We report hereafter the main design choices and results for each of these
studies and experiments. The interested reader can find supplementary details
in~\cite{jacq2016building, hood2015when}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{System Validation at Schools}
\label{studies_school}

Over the two years of the project, we conducted four studies in schools
(Figure~\ref{fig:schools}). These experiments were meant to technically validate
the system (is it actually able to autonomously write and learn from handwriting
demonstrations?) and test the interaction (is the apparatus easy to grasp and to
interact with for children?). We also studied the initial acceptance of the
robot in the school environment (through several formal and informal discussions
with teachers) and how children engage with the robot (and maintain or not this
engagement).

\begin{figure}
    \centering
    \includegraphics[height=5cm]{schools2}
    \caption{\small The first field studies were focused on technical validation, with
    more than 70 pupils interacting with the robot over short periods (between
    10 and 25 minutes), either alone or in small groups.}
    \label{fig:schools}
\end{figure}

Critically, these studies were conducted with \emph{whole classes}: we decided
not to select specifically underperforming children as having more children (73
in total) was beneficial for these preliminary studies. In addition, due to ethical
concerns, this would have required complex organization with the school that we
wanted to avoid at the validation stage.

\paragraph{System validation}

From a technical perspective, the system achieves an acceptable level of
reliability and allows a technically sound autonomous interaction. For instance,
during the second school study (\textit{School B}), the robot withstood
interactions which lasted for a total of 160 minutes.  During this time the
robot wrote 335 letters, 152 of which in response to demonstrations received
from the 21 children. Technical intervention was only required for the three
instances that the robot fell during that day.  Otherwise, the technical
components of the system operated autonomously and as expected with all the
groups of children.

Due to the modular software architecture (mostly independent ROS nodes), the
occasional crashes occurring during others studies were usually quickly resolved
by re-launching the faulty node alone, and did not significantly impact the
interaction.

The otherwise technical limitations were related to some letters or writing
styles (most notably, the ones requiring multiple strokes per letter) not being
adequately processed by the learning algorithm. Support for such letters was
added as a follow-up to the validation studies.

\paragraph{Acceptance}

Children's recognition that the robot is writing by itself is critical for our
approach to be effective. When asked, no child indicated that they did not
believe that the robot was writing by itself. There were, at times, questions
about the robot's writing without a pen at the beginning of the interaction, but when
advised that the robot ``tells the tablet what it wants to write,'' this was
accepted by the children.

During two focus groups organized with teachers, they welcomed the approach and
recognized it as useful and promising; this was \textit{a posteriori} confirmed
by multiple spontaneous contacts made by parents and therapists who were looking
forward to using the system with their children.

\paragraph{Sustained Engagement}

Factually, the children engaged into the teaching activity: in the
study at school B, for instance, they demonstrated an average of 10.9
demonstration letters (SD = 4.4) for an average session duration of 11 minutes.
In 9 out of the 14 sessions (64\%), the robot received demonstration letters
even \emph{after} reaching the final stage of the interaction, suggesting an
intrinsic motivation to further engage in the interaction.

%\begin{figure}[ht!]
%    \centering
%    \includegraphics[width=0.9\linewidth]{engagement}
%    \caption{\small Engagement level of one child, manually annotated by two judges
%        over a 13 min long interaction (Cronbach’s alpha test: $\alpha = .82$,
%        good inter-judge agreement). 0 means ``complete disengagement from the
%        task'', 6 means ``full engagement''. The vertical lines represent activity
%    switches between handwriting teaching and story-telling (used as a
%    disturber).}
%    \label{engagement_level}
%\end{figure}

We also conducted a quantitative assessment of the engagement levels of the
children. Table~\ref{tab:results-with-me-ness} reports the levels of
\emph{with-me-ness} of the children during the second study at school A.
\emph{With-me-ness} is a quantifiable precursor of engagement: it measures the
percentage of time spent by the child focusing on the task at hand~\cite{lemaignan2016realtime}. We compute it by first estimating the focus of
attention of the child over the course of the interaction (using real-time 6D
head pose estimation); then, by matching this measured focus of attention
with the \emph{expected} attentional targets. For instance, when the child is supposed
to write a demonstration, we expect her to look at the tablet if she is actually
engaged into the task; when the robot is telling a story,
we expect the engaged child to look at the robot, etc.

\begin{table*}[h!]
    \centering
    \caption{\textbf{Levels of with-me-ness}. Percentage of interaction time
        during which the child was effectively focusing her attention on the
        task. The six children are those from the second study at School A.
        Interaction duration: M = 19.6 min, SD = 1.58. Results
        taken from~\cite{lemaignan2016realtime}.}

    \begin{tabular}{p{1cm}cccccccc}
        \toprule
        \bf Child & 1 & 2 & 3 & 4 & 5 & 6 & {\bf M} & {\it SD} \\
        \midrule
        $\mathcal{W}$ & 79.4\% & 81.6\%  & 90.5\% & 87.9\% & 90.7\% & 80.9\% & {\bf 85.2\%} & {\it 5.1} \\ 
        \bottomrule
    \end{tabular}
    \label{tab:results-with-me-ness}
\end{table*}

The average with-me-ness is well above 80\%, and confirms that the children
were very much engaged into these 20 minutes of interaction with the robot,
paying close attention to the task.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Clinic Study: How Children Take on the Role of a Teacher}\label{normandie}
\label{study_normandy}

\paragraph{Context, Study Design}

This experiment was conducted to study how easily children with actual
deficits take on the role of a teacher, and to what extent they adopt this role
(measuring actual handwriting improvements was not a primary goal of this
experiment).

The experiment took place at an occupational therapy clinic in Normandy, France.
Eight children participated, selected by the occupational therapist based on
their age and type of deficit (all related to handwriting).  Given their age (6
to 8 years old) and school year, all of these children would be expected to know
how to correctly shape cursive letters. Each child attended three times a one
hour long session spread over two weeks (except for two of them who only
attended one session). The experimenter's role was limited to the explanation of
the task and the basic tablet usage. The children were provided with two
tablets: one to choose the words (or letters) to teach to the robot, one to
write.

We only provided the children with minimal explanations on the task (they would
have to help the robot to improve its writing style), so as to assess if and how
children would naturally take on the role of the teacher. We additionally
assessed how seriously they engage into helping the robot through two additional
buttons on the tablet: a green ``thumbs up'' and a red ``thumbs down''. The
children were told to freely use them to evaluate the robot's improvements
(``thumb up'' to give positive feedback, ``thumb down'' to give negative
feedback). Our assessment builds on the hypothesis that the more the child
provides feedback to the robot, the more they assume the role of a teacher.
Then, by correlating the feedback with the actual performance of the robot, we
can measure to what extent the children are adopting their teacher role: if
their feedback does correlate with the actual performance of the robot, the
child has likely successfully adopted the teacher role.

\paragraph{Results}

All children maintained their engagement during the whole sessions. They
provided on average 42 demonstrations per session. All children used the
feedback buttons (in total, 99 ``thumbs up'' and 33 ``thumbs down'' were
recorded). This indicates that they are all able to engage into playing the role
of a teacher.

To study the correlation between the children's feedback and the actual
improvements of the robot, we estimate the robot's progress as the difference
between an initial score (euclidian distance between the shape drawn by the
robot at its first attempt and the shape of the reference letter) and the
robot's score at the current round of demonstrations. We then correlate this
progress to the positive or negative feedback provided by the children (details
of the method, as well as the complete results, are presented
in~\cite{jacq2016building}).

We find that the feedback of five out of the eight children does significantly
correlate with the actual performance of the robot.  This indicates that these
children are effectively taking on the teacher's role and are seriously
providing feedback to the robot. The observation of the three remaining children
reveal a variety of behaviors (for instance, one was actually rating how
``nice'' the robot is, and another one had rather ``artistic'' writing style
preferences that were independent of the actual legibility of the robot), but
none of these children adopted a playful-only behavior toward the robot.

To summarize, this experiment shows that children with actual handwriting
impairments do accept well the robot, can commit themselves into long
interactions, and do adopt the role of the teacher (as evidenced by both the
number of writing demonstrations they provided and their self-inclination to
give adequate feedback to the robot).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Case Study 1: Vincent}
\label{vincent}

For the first case study, we invited and followed Vincent, a six year old
child, once per week over a period of a month. Our primary aim was to address
the question of whether we can sustain Vincent's engagement and commitment to
the writing activity over such a period.

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{diego}
    \caption{\small Vincent correcting {\sc nao}'s attempt by rewriting the
        whole word. Empty boxes are drawn on the screen to serve as a template for the child
        and to make letter segmentation more robust.}
    \label{fig:diego}
\end{figure}

The study took place at our laboratory (Figure~\ref{fig:diego}), and we chose to
design the activity around a storyline meant to be attractive for a 6 year old
boy: one of our {\sc nao}s was away for a mysterious scientific mission, and it
needed the support of another one -- which would remain at the lab -- to
interpret curious pictures that were sent every week. Vincent had to help the
second robot understanding the pictures, and since the two robots had somehow
beforehand agreed to communicate with ``letters written like humans'' (\ie
handwritten), Vincent also had to help the robot to write good-looking letters
(because, well, this robot was terrible at writing!). The experimental setup was
similar to Figure~\ref{experimental_setup}, except that Vincent had to tell the
robot what to write with small plastic letters (visible behind the robot on
Figure~\ref{fig:diego}).

To supplement the intrinsic motivation of helping a robot to communicate with
another one, we gradually increased the complexity of Vincent's task to keep it
challenging and interesting (the first week: demonstration of single letters;
the second week: short words; the third week: a full letter --
Figure~\ref{fig:stimuli}).

The last session was set as a test: the ``explorer'' robot had come back from
its mission and it actually challenged the other robot in front of Vincent:
\emph{``I don't believe you wrote yourself these nice letters that I received!
Prove it to me by writing something in front of me!''} This situation was meant
to evidence the Protégé effect: by judging the other robot's handwriting, the
``explorer'' robot would implicitly judge Vincent's skills as teacher, and in
turn, Vincent's handwriting.

\paragraph{Results}

\begin{figure}
    \centering
    \subfigure[Initial text, generated by the robot]{
        \includegraphics[height=6cm]{diego-initial-letter}
    }
    \subfigure[Final text, after training with Vincent]{
        \includegraphics[height=6cm]{diego-final-letter}
    }

    \caption{\small Text (in French) generated by the robot, before and after a one
        hour long interaction session with the child. The red box
        highlights one instance of striking improvement of the robot's
        handwriting legibility \emph{`envoyer'}.}

    \label{fig:stimuli}
\end{figure}

Over the whole duration of the study, Vincent provided 154 demonstrations to
the robot, and he remained actively engaged over the four weeks. The story was
well accepted by the child and he seriously engaged into the game. After the
first week, he showed good confidence to play with the robot and by the end of
the study he had built affective bonds with the robot, as evidenced by several
letters he did send to the robot \emph{after} the end of the study (one of them
four months later) to get news. This represents an initial validation of our
hypothesis: our system can effectively keep a child engaged with the robot for a
relatively long period of time (about 5 hours spread over a month), and we can
build a tutor/protégé relationship.

No hard conclusion can be drawn in terms of actual handwriting remediation as we
did not design this study to formally assess possible improvements. However, as
visible on Figure~\ref{fig:stimuli}, Vincent was able to significantly improve
the robot's skill, and  he acknowledged that he was the one helping the robot
during post-hoc interviews. In that respect, Vincent realized that he was ``good
enough'' at writing to help someone else. The fact that Vincent adopted
the role of a teacher is further supported by feedback sent by
Vincent's parents a week after the end of the experiment: ``Vincent's
handwriting has changed over the last weeks, going from a mix of standalone and
cursive letters to full words in cursive. This requires a lot of effort and
concentration from him, but he did succeed during the sessions with the robot as
he knew he had to show a consistent style of writing to the robot''.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Case Study 2: Thomas}
\label{thomas}

The second long-term study was designed in collaboration with an occupational
therapist in Geneva, and aimed at deploying the system on the longer-term with a
real therapeutic case.

Thomas is a 5.5 year old child. He has been diagnosed with visuo-constructive
deficits, which translate into difficulties for him to consistently draw
letters. In addition, focusing on a task is difficult for Thomas, who tends to
rapidly shift his attention to other things. Since the robot's learning
algorithm requires repeated demonstrations of similarly shaped letters to
converge, the occupational therapist was especially interested in observing if
the robot would induce a strong enough motivation for Thomas to focus on
producing many regular, consistent letters, thus overcoming his deficit.

The experiment took place at the therapist's clinic (four sessions spread
over 5 weeks). Contrary to Vincent's experiment, we chose not to introduce any
backstory beyond a simple prompt (``the robot wants to participate in a robotic
handwriting contest, will you help him prepare?'') only provided during the
first session. Hence, we also tested during this case-study if the robot (and
the Protégé effect) would induce by itself a strong enough intrinsic motivation
to keep the child engaged over the five weeks.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{learning_6_demos}

    \vspace{1em}

    \includegraphics[width=0.7\linewidth]{learning_6_distances}
    \caption{\small Demonstrations provided by Thomas for the number ``6'' (top
        row) and corresponding shapes generated by the robot. The plot beneath
        shows the distance to the reference shape (in the eigenspace of the
        shape) to Thomas' demonstrations, and to the robot's attempts. After
        eight demonstrations, Thomas decided that the robot's ``6'' was good
        enough, and switched to another character. In that respect, he was the
        one leading the learning process of the robot.}

    \label{learning_6_demos}
\end{figure}

The occupational therapist had recently carried out activities with Thomas on
writing numbers, so we decided with her to focus on these as well: Thomas would
use a secondary tablet to tell the robot what number to write, and would then
correct the robot's attempts like in the other experiments.
Figure~\ref{learning_6_demos} shows the attempts/corrections cycles that
occurred during one of the session, on the number ``6''.

%Before the experiment, Thomas was working on writing numbers with the therapist.
%Hence we decided to turn the CoWriter activity to teach numbers to the robot.
%Build on existing iPad applications actually used by occupational therapists
%(Dawson Toth's \emph{ABC's Writer}), we designed an Android application for
%pre-test and post-test. It consisted in drawing numbers following an helping
%pattern, that becomes more fine with levels in order to increase
%difficulty~\ref{fig:abc-writer}.

Since Thomas would frequently draw mirrored numbers, or hard-to-recognize shapes
(see Figure~\ref{Thomas_progress}), the learning algorithm of the robot
initially tended to converge towards meaningless scribblings. We addressed this
issue by having the robot \emph{refuse} allographs that were too far from the
reference shape (the robot would instead say \emph{``I'm not sure I understand
what you are drawing...''}), so that the child had to pay good attention to what
he would demonstrate to the robot. Also, to make the robot's progress evident,
we modified the initialization step of the learning algorithm to start with a
roughly vertical stroke instead of a deformed number (see the initial state on
Figure~\ref{learning_6_demos}).

%\begin{figure}
%    \centering
%    \includegraphics[width=0.9\linewidth]{abc-writer}
%    \caption{\small Screenshot of the Android application developed to be used as
%    pre-test and post-test: the child must follow with his finger the path of
%the letter or number. We count the number of times the finger goes outside of
%the red-bordered area.}
%    \label{fig:abc-writer}
%\end{figure}
%

\paragraph{Results}

Despite his attention deficit, Thomas was able to remain engaged in the activity
during more than forty minutes in each session (a long time for a five years
old). In total, 55 allographs out of 82 demonstrated by the child were
acceptable considering our threshold (with a progressive improvement from 13 out
of 28 in the first session up to 26 out of 29 in the last session).

As soon as Thomas understood that the robot was only accepting well-formed
allographs, he started to focus on it and he would typically draw 5 or 6 times
the number before actually sending to the robot (the tablet lets children clear
their drawing and try again before sending it). According to the therapist, it
was the first time that Thomas was seen to correct himself in such a way,
explicitly having to reflect on how \emph{another agent} (the robot) would
interpret and understand his writing. Figure~\ref{Thomas_progress} shows how he
gradually improved his demonstrations for two different numbers.

Since the robot's handwriting started from a simple primitive (a stroke), each
time Thomas succeeded in having his demonstration accepted by the robot, the
improvement was clearly visible (as shown in Figure~\ref{learning_6_demos}).
This led to a self-rewarding situation that effectively supported Thomas'
engagement.

\begin{figure}
    \centering
    \subfigure[Number ``2'']{
        \includegraphics[width=0.45\linewidth]{henry2}
    }
	\subfigure[Number ``5'']{
        \includegraphics[width=0.45\linewidth]{henry5}
    }
    \caption{\small Normalized distance between Thomas' demonstrations and
        reference allographs for the numbers ``2'' and ``5''. The horizontal
        dashed line correspond to the threshold for the robot to accept a
        demonstration. Thomas' progress is visible on these figures: we find a
        significant negative regression equation ($r=-0.023, F(1,19)=8.69,
        p<.02$, adjusted $R^2=.461$) for the number ``2'' (dotted red line),
        indicating that Thomas' shapes are getting closer to the reference. The
        regression is not significant for the number ``5'', but we can observe
        that after about 10 repetitions, all the demonstrations are deemed of
        acceptable quality by the robot.  }

    \label{Thomas_progress}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary of the Findings and Discussion}

The four school studies, the experiment at the clinic and the two case-studies
provide a first broad picture of how a robot-based remediation to handwriting is
accepted by the children and practitioners, and what outcomes can be expected.

In total, more than 80 children have interacted with the system, for a total
duration of more than 38 hours, in multiple experimental configurations
(individual interaction \vs pairs \vs groups; at school, in the lab, at an
occupational therapy clinic; short interactions \vs long, repeated interactions;
5 to 8 years old
children). We summarize hereafter the main findings from these experiments and
discuss some of the critical points of our approach.

\paragraph{Technical Assessment} We set ourselves the challenge of developing an
autonomous robotic system able to perform handwriting tasks with children in
real world environments. The system performed generally well: the children had no
issues interacting with the robot, the robot experienced relatively few crashes,
and those crashes did not significantly impair the interactions. The system
has furthermore proved robust enough to conduct several hour-long experiments.
However, we also found that the level of expertise required to deploy and
operate our system still makes the presence of an experimenter mandatory at all
time. Non-trivial technical developments may be required to reach a level of
usability suitable for a broader, non-expert, audience.

\paragraph{Acceptance} The acceptance of the
robot in a educative and/or medical context was not initially
obvious, and indeed, the attitude of practitioners toward the robot varies.
For instance, while Thomas' therapist did
voluntarily contact us (after hearing about the project on the radio) and
readily involved herself in the design of the experiment, we had
less positive feedback from other practitioners: in one instance, we contacted
ourselves a local group of school psychologists to
present the project: after the meeting, the opinions were definitively mixed,
with some therapists willing to conduct actual experiments with their children,
and others not quite as enthusiastic.

Expectedly the reaction of the children to the robot was good: they enjoyed
interacting with the system, and, as we have shown, they actually committed to
their teacher role over extended periods of time. The teaching situation was
well accepted and technical constraints like having the robot to only gesture
writing on the tablet instead of actually \emph{physically} writing with a pen,
were not actually raised as issues.

\paragraph{Engagement} We measured the engagement of the children by three
different means: the number of demonstrations they provided to the robot, the
amount of qualitative feedback they gave to the robot (in the \emph{Clinic}
experiment), and to what extent they were focusing on the task (measure of the
with-me-ness). Independently of the experimental setting, it appears that the
children engage easily into the interaction. More interestingly, we show in the
\emph{Clinic} experiment that they generally take on the role of a teacher
easily, that they act this role seriously (and not only playfully), and that
they can assume this role over an extended period of time. Designing a system
that keeps children engaged over several hours is especially important for
handwriting remediation, and we show in the case-studies that even children as
young as 5.5 years old like Thomas were able to do so. This seems to indicate
that our approach could be relevant for a broad range of ages.

Another finding relates to the Protégé effect: as seen with Vincent (when he
sends us a mail several months later to know how the robot is doing with its
writing) or with Thomas (when he realizes that the robot ``does not
understand'', and consequently quickly improves his own writing to better help
the robot), our system does elicit a Protégé effect that not only helps the
children to remain engaged in their teaching over several weeks, but also
positively impacts their learning process (Vincent strives to write in a
consistent manner, as does Thomas). This supports {\it a posteriori} our choice
to build an interaction situation based on the \emph{learning by teaching}
paradigm.

\paragraph{Remediation Efficacy} Vincent's case study did provide us with
initial material to evidence handwriting improvements (Figure~\ref{fig:stimuli})
and the study with Thomas provides further data, both quantitative
(Figure~\ref{Thomas_progress}) and qualitative (feedback from the therapist that
points how Thomas is much better at drawing consistent shapes in a repeated manner,
as well as reflecting on his own performance by training several times before
actually sending a demonstration to the robot).

We must however remain cautious here as for the actual role of the system: while
the children were indeed the ones deciding what to teach to the robot and the
robot was autonomously learning and responding, the role of the adults (the
experimenters or the occupational therapists) should not be underestimated.
Beyond the normal explanations of how to operate the tablet and how to interact
with the robot, the adults played the role of a facilitator in each of the
studies by prompting the children to comment on the robot's performance,
suggesting possible corrections, or proposing to try another letter/number/word.
This facilitation not only compensates for the possible shortcomings of the interaction,
but is also a fundamental part of the learning process itself. In that respect,
our robot is essentially a tool that creates a favorable learning situation for
the child, and where the adult (be it a teacher or a therapist) keeps their entire
educative role.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

We believe that this research provides a novel perspective on educative robots
at several levels. Namely, we show that:

\begin{itemize}
    \item robots in an educative context are certainly \textbf{relevant and
        effective beyond STEM} (Science, Technology, Engineering and
        Mathematics) \textbf{teaching},

    \item we can successfully \textbf{transpose the well-established
        \emph{learning by teaching} paradigm} from education sciences to
        robotics, even in a \textbf{complex form}: handwriting is a difficult
        physical skill, the robot learns and interact autonomously, the child is
        responsible not only for the teaching but also for the teaching
        orchestration by managing the turn taking and the progression of the
        activity,

    \item blending machine-learning techniques with human-robot interaction
        allows for building a \textbf{believable agent}, that induces
        \textbf{social commitment},

    \item this social commitment induces \textbf{cognitive engagement}
        of the child with the robot, which is \textbf{a key learning lever} as it
        elicits \textbf{reflective, meta-cognitive mechanisms} on the learning task, and that

    \item we have been able to \textbf{sustain a long-term interaction}
        (several hours) involving a task that would typically be considered 
        repetitive and challenging by the children.

\end{itemize}

Supported by several field experiments, our initial results are promising as we
ultimately evidence handwriting improvements. It is however still early to quantify the
lasting effects of this remediation: handwriting is a complex cognitive skill,
that builds on many individual and social factors. Self-confidence is one of
them. Our approach endows the child with the role of a teacher who can help a
robot: we expect it may as well help some children to recover self-esteem and
self-confidence by putting them in a positive, gratifying role. The experiments
that we have conducted so far do not allow us to confirm this hypothesis yet,
and more research will have to be conducted in this direction.

\paragraph{Possible Ethical Concerns} There are two aspects of this research ought to be
discussed in terms of their possible ethical implications: the perceived role of
the robot \textit{vis-à-vis} the real teachers, and the implications of the
mentor-protégé relationship for children, especially vulnerable ones.

The place and role of the robot \textit{vis-à-vis} the teacher can be
questioned: as we see it, the role of the robot within the classroom (or at the
therapist's clinic) does not infringe upon the role of the adult (teacher or
therapist). The core of the \emph{learning by teaching} paradigm relies on the
child becoming the teacher of an underperforming pupil (the robot): from that
perspective, the robot does not replace the teacher, on the contrary. It plays a
different role in the classroom, which happens to be novel as well: the robot is
the \emph{least} performing student, and still a very patient, always eager to
improve, one. In our experience, the teacher/facilitator keeps an instrumental
role during the interaction, and the learning would hardly occur if the child is
left alone (or even semi-alone). The initial feedback that we received from the
teachers during the focus groups confirms this perception: once explained, our
approach make sense to them and is indeed welcomed as a relevant pedagogical
innovation.

The implication of the mentor-protégé relationship on the children is less
clearly understood. We have certainly seen that the children can establish
strong affective bonds with the robot (as witnessed for instance by the letter
sent by Vincent several months after he interacted with the robot), but we are
not yet able to precisely characterize these bonds. The ethical implications of
the mentor-protégé relationship have been explored before in the context of
human teaching~\cite{brad1999mentor,wendelyn2008context}, but they mostly looked
at the question from the perspective of the protégé, whereas in our case, the
child is the mentor. As such, relatively little is known on the psychological
implications for a child to commit to helping a robot, and as advocated by
Belpaeme and Morse~\cite{belpaeme2010time}, we likely need to first gain more
field experience before being able to draw conclusions.

Beyond handwriting, we do however believe that this work provides a novel
perspective on the role for robots in the field of education. \emph{Learning by
teaching} is a powerful paradigm because of not only its pedagogical efficacy,
but its potential to positively impact the child's motivation and self-esteem.
While we need to carefully clear up the possible ethical concerns, we hope that
this article shows that this is a very relevant context of use for robots: when
facing a child with school difficulties, robots can play the role of a naïve
learner which neither adults nor peers can convincingly play.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgments}

We warmly thank all the children who participated to the studies, and the
teachers and staff of the International School of Geneva, the Institut Florimont
and the Haut-Lac school for their support. We express our particular gratitude
to Anne Gossin and Fanny Lebel for inviting us to their clinics.

This research was partially supported by the Funda\c{c}\~{a}o para a Ci\^{e}ncia
e a Tecnologia (FCT) with reference UID/CEC/50021/2013, and by the Swiss
National Science Foundation through the National Centre of Competence in
Research Robotics.

\bibliographystyle{abbrv}
\bibliography{biblio}

\end{document}